---
title: "hw7"
author: "Andrew Chen"
date: "2023-03-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(tidyverse)
```

```{r}
#data(happy, package="productplots") #not used
HAPPY <- readRDS("data/HAPPY.rds")
#head(HAPPY)
HAPPY1 <- HAPPY
```
PART 1
1. Data cleaning: the values “IAP”, “DK” and “NA” all encode missing values. Replace all of these instances by the value NA.

```{r}
#Check for DK, IAP, NA
#HAPPY, AGE, SEX, MARITAL, DEGREE, FINRELA, HEALTH, WTSSALL, PARTYID, POLVIEWS
tempText = HAPPY1$HAPPY
tempDK = HAPPY1 %>% filter( tempText== "DK")
tempIAP = HAPPY1 %>% filter(tempText == "IAP")
tempNA = HAPPY1 %>% filter(tempText == "NA")
```

```{r}
#Replace all IAP, DK, NA, with NA
HAPPY1[HAPPY1 == "DK"] <- NA
HAPPY1[HAPPY1 == "IAP"] <- NA
HAPPY1[HAPPY1 == "NA"] <- NA
```

2. Check the type of the variable and cast into the right type (factor variable for categorical variables). For age, change “89 OR OLDER” to 89 and assume the variable should be numeric.
3. Bring all levels of factors into a sensible order. For marital you could e.g. order the levels according to average age.

HAPPY-factor
```{r}
HAPPY1$HAPPY = factor(HAPPY1$HAPPY, 
                      levels = c("NOT TOO HAPPY", "PRETTY HAPPY", "VERY HAPPY"))
```

AGE-cast
```{r}
HAPPY1[HAPPY1 == "89 OR OLDER"] <- 89
HAPPY1$AGE = as.numeric(HAPPY1$AGE, na.rm=TRUE)
```

SEX-factor
```{r}
HAPPY1$SEX = factor(HAPPY1$SEX, levels = c("MALE", "FEMALE"))
```

MARITAL-factor
```{r}
HAPPY1$MARITAL = factor(HAPPY1$MARITAL, 
                        levels = c("NEVER MARRIED", "MARRIED", "DIVORCED"))
```

DEGREE-factor
```{r}
HAPPY1$DEGREE = factor(HAPPY1$DEGREE, 
                       levels = c("LT HIGH SCHOOL", "HIGH SCHOOL", 
                                  "JUNIOR COLLEGE", "BACHELOR", "GRADUATE"))
```

FINRELA-factor
```{r}
HAPPY1$FINRELA = factor(HAPPY1$FINRELA, 
                        levels = c("FAR BELOW AVERAGE", "BELOW AVERAGE", 
                                   "AVERAGE", "ABOVE AVERAGE", "FAR ABOVE AVERAGE"))
```

HEALTH-factor
```{r}
HAPPY1$HEALTH = factor(HAPPY1$HEALTH, levels = c("POOR", "FAIR", "GOOD", "EXCELLENT"))
```

WTSSALL-cast
```{r}
HAPPY1$WTSSALL = as.numeric(HAPPY1$WTSSALL, na.rm=TRUE)
```

PARTYID-factor
```{r}
HAPPY1$PARTYID= factor(HAPPY1$PARTYID, 
                       levels = c("OTHER PARTY","STRONG DEMOCRAT", "NOT STRONG DEMOCRAT",
                                  "IND, NEAR DEM", "INDEPENDENT", "IND, NEAR REP", 
                                  "NOT STR REPUBLICAN", "STRONG REPUBLICAN"))
```

POLVIEWS-factor
```{r}
HAPPY1$POLVIEWS = factor(HAPPY1$POLVIEWS, 
                         levels = c("EXTREMELY LIBERAL", "LIBERAL", 
                                    "SLIGHTLY LIBERAL", "MODERATE", 
                                    "SLGHTLY CONSERVATIVE", "CONSERVATIVE", 
                                    "EXTRMLY CONSERVATIVE"))
```

PART 2


Happiness by age and sex.
```{r}
HAPPY1 %>% ggplot(aes(x = AGE, y = SEX, fill = HAPPY)) + 
  geom_boxplot() +
  facet_wrap(~HAPPY)
```
Not sure why NA is still showing on graph. Most people are between the ages of 30-60. The average "VERY HAPPY" age for men and women is around 46. The average "PRETTY HAPPY" age is higher for women than men. Same for "NOT TOO HAPPY". 

