---
title: "hw7"
author: "Andrew Chen"
date: "2023-03-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(dplyr)
library(tidyverse)
data(happy, package="productplots")
HAPPY <- readRDS("data/HAPPY.rds")
#head(HAPPY)
HAPPY1 = HAPPY
```

1. Data cleaning: the values “IAP”, “DK” and “NA” all encode missing values. Replace all of these instances by the value NA.

```{r}
#Finding all the locations of missing values
#HAPPY, YEAR, AGE, SEX, MARITAL, DEGREE, FINRELA, HEALTH, WTSSALL, PARTYID, POLVIEWS
tempText = HAPPY1$POLVIEWS

tempIAP = HAPPY1 %>% filter(tempText == "IAP")
#FINRELA-4383
#HEALTH-16325
#PLVIEWS-6777

tempDK = HAPPY1 %>% filter(tempText == "DK") 
#AGE-1
#DEGREE-30
#FINRELA-468
#HEALTH-30
#PARTYID-11
#POLVIEWS-2247

tempNA = HAPPY1 %>% filter(tempText == "NA")
#none
```

```{r}
HAPPY2 <- HAPPY1 %>% 
  mutate(FINRELA = str_replace(FINRELA, "IAP", "NA"))

```

2. Check the type of the variable and cast into the right type (factor variable for categorical variables). For age, change “89 OR OLDER” to 89 and assume the variable should be numeric.
```{r}

```

3. Bring all levels of factors into a sensible order. For marital you could e.g. order the levels according to average age.
```{r}

```

