---
title: "hw7"
author: "Andrew Chen"
date: "2023-03-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(tidyverse)
```

```{r}
#data(happy, package="productplots") #not used
HAPPY <- readRDS("data/HAPPY.rds")
#head(HAPPY)
HAPPY1 <- HAPPY
```
PART 1
1. Data cleaning: the values “IAP”, “DK” and “NA” all encode missing values. Replace all of these instances by the value NA.

```{r}
#Check for DK, IAP, NA
#HAPPY, AGE, SEX, MARITAL, DEGREE, FINRELA, HEALTH, WTSSALL, PARTYID, POLVIEWS
tempText = HAPPY1$HAPPY
tempDK = HAPPY1 %>% filter( tempText== "DK")
tempIAP = HAPPY1 %>% filter(tempText == "IAP")
tempNA = HAPPY1 %>% filter(tempText == "NA")
```

```{r}
#Replace all IAP, DK, NA, with NA
HAPPY1[HAPPY1 == "DK"] <- NA
HAPPY1[HAPPY1 == "IAP"] <- NA
HAPPY1[HAPPY1 == "NA"] <- NA
```

2. Check the type of the variable and cast into the right type (factor variable for categorical variables). For age, change “89 OR OLDER” to 89 and assume the variable should be numeric.
3. Bring all levels of factors into a sensible order. For marital you could e.g. order the levels according to average age.

HAPPY-factor
```{r}
HAPPY1$HAPPY = factor(HAPPY1$HAPPY, 
                      levels = c("NOT TOO HAPPY", "PRETTY HAPPY", "VERY HAPPY"))
```

AGE-cast
```{r}
HAPPY1[HAPPY1 == "89 OR OLDER"] <- 89
HAPPY1$AGE = as.numeric(HAPPY1$AGE, na.rm=TRUE)
```

SEX-factor
```{r}
HAPPY1$SEX = factor(HAPPY1$SEX, levels = c("MALE", "FEMALE"))
```

MARITAL-factor
```{r}
HAPPY1$MARITAL = factor(HAPPY1$MARITAL, 
                        levels = c("NEVER MARRIED", "MARRIED", "DIVORCED"))
```

DEGREE-factor
```{r}
HAPPY1$DEGREE = factor(HAPPY1$DEGREE, 
                       levels = c("LT HIGH SCHOOL", "HIGH SCHOOL", 
                                  "JUNIOR COLLEGE", "BACHELOR", "GRADUATE"))
```

FINRELA-factor
```{r}
HAPPY1$FINRELA = factor(HAPPY1$FINRELA, 
                        levels = c("FAR BELOW AVERAGE", "BELOW AVERAGE", 
                                   "AVERAGE", "ABOVE AVERAGE", "FAR ABOVE AVERAGE"))
```

HEALTH-factor
```{r}
HAPPY1$HEALTH = factor(HAPPY1$HEALTH, levels = c("POOR", "FAIR", "GOOD", "EXCELLENT"))
```

WTSSALL-cast
```{r}
HAPPY1$WTSSALL = as.numeric(HAPPY1$WTSSALL, na.rm=TRUE)
```

PARTYID-factor
```{r}
HAPPY1$PARTYID= factor(HAPPY1$PARTYID, 
                       levels = c("OTHER PARTY","STRONG DEMOCRAT", "NOT STRONG DEMOCRAT",
                                  "IND, NEAR DEM", "INDEPENDENT", "IND, NEAR REP", 
                                  "NOT STR REPUBLICAN", "STRONG REPUBLICAN"))
```

POLVIEWS-factor
```{r}
HAPPY1$POLVIEWS = factor(HAPPY1$POLVIEWS, 
                         levels = c("EXTREMELY LIBERAL", "LIBERAL", 
                                    "SLIGHTLY LIBERAL", "MODERATE", 
                                    "SLGHTLY CONSERVATIVE", "CONSERVATIVE", 
                                    "EXTRMLY CONSERVATIVE"))
```

PART 2
Repo link: https://github.com/wardennnnnn/ds202-hw7

Kennedy Wendl 

```{r}

ggplot(HAPPY1, aes(x = HAPPY, y = AGE)) +
  geom_boxplot() + 
  facet_grid(HEALTH ~ DEGREE)

```
Question: Does health and degree status have an affect on level of happiness?

To answer this question, I created a boxplot which includes the age, level of happiness, degree, and health of the individuals. As shown on the plot above, people with poor health and less than high school level of education are generally more unhappy than those with good/excellent health and who have graduated high school. The uppermost graph on the left highlights this fact, as the level of happiness increases the group size decreases. That being said, this relationship isn't particularly strong. As education and health increases, it becomes harder to determine if they have a substantial effect. Additionally, this could also be a result of varying age groups. 


Vrindha Vegiraju
```{r}

ggplot(HAPPY1, aes(x = HAPPY, y = 89)) +
  geom_boxplot() + 
  facet_grid(~HEALTH)

```
For this question, I changes the age variable to 89 for it to be numeric. I used the age variable to create a boxplot and find the correlation between happiness and a person's age. The graph is not very strong but shows that there are fewer people as happiness increases. Because the ages are so large, the results may not be very clear.


Andrew Chen

Happiness by age and sex.
```{r}
HAPPY1 %>% ggplot(aes(x = AGE, y = SEX, fill = HAPPY)) + 
  geom_boxplot() +
  facet_wrap(~HAPPY)
```
Most people are between the ages of 30-60. The average "VERY HAPPY" age for men and women is around 46. The average "PRETTY HAPPY" age is higher for women than men. Same for "NOT TOO HAPPY". 

