---
title: "hw7"
author: "Andrew Chen"
date: "2023-03-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(tidyverse)
```

```{r}
#data(happy, package="productplots") #not used
HAPPY <- readRDS("data/HAPPY.rds")
#head(HAPPY)
HAPPY1 <- HAPPY
```

1. Data cleaning: the values “IAP”, “DK” and “NA” all encode missing values. Replace all of these instances by the value NA.

```{r}
#Check for DK, IAP, NA
#HAPPY, AGE, SEX, MARITAL, DEGREE, FINRELA, HEALTH, WTSSALL, PARTYID, POLVIEWS
tempText = HAPPY1$FINRELA
tempDK = HAPPY1 %>% filter( tempText== "DK")
tempIAP = HAPPY1 %>% filter(tempText == "IAP")
tempNA = HAPPY1 %>% filter(tempText == "NA")
```

```{r}
#Replace all IAP, DK, NA, with NA
#https://www.statology.org/r-replace-values-in-data-frame-conditionally/
HAPPY1[HAPPY1 == "DK"] <- NA
HAPPY1[HAPPY1 == "IAP"] <- NA
HAPPY1[HAPPY1 == "NA"] <- NA
```


2. Check the type of the variable and cast into the right type (factor variable for categorical variables). For age, change “89 OR OLDER” to 89 and assume the variable should be numeric.
```{r}
HAPPY2 = HAPPY1
```

HAPPY-factor
```{r}
HAPPY2 %>% factor(HAPPY)
```
HAPPY-cast
AGE-cast
SEX-factor
MARITAL-factor
DEGREE-factor
FINRELA-factor
HEALTH-factor
WTSSALL-cast
PARTYID-factor
POLVIEWS-factor


3. Bring all levels of factors into a sensible order. For marital you could e.g. order the levels according to average age.
```{r}

```

